### 2022-01-27 Paper/SupMat.Rnw
draft of Discussion done

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xr}
\usepackage{authblk}
\usepackage{lineno}

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}

%new commands and so on
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcB}{\mathcal{B}}
\newcommand{\Rp}{\text{Re}}
\newcommand{\var}{\text{var}}
\newcommand{\cov}{\text{cov}}
\newcommand{\cor}{\text{cor}}
\let\conj\overline

%set up the theorem environment
\newtheorem{thm}{Theorem}
\newtheorem{corl}[thm]{Corollary}

%external documents
\externaldocument[MT-]{Paper}

%Get line numbers to work, also in figure captions
\linenumbers
\DeclareCaptionFont{linenumbers}{\internallinenumbers}
\captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption

%header material for paper
\title{Supporting information for ``How environmental drivers of synchrony interact''}
\date{}

%***Tentative author order, Bell last for his data contributions
\author[a,*]{Daniel C. Reuman}
\author[b]{Max Castorani}
\author[c]{Kyle Cavanaugh}
\author[d]{Lawrence W. Sheppard}
\author[b]{Jonathan Walter}
\author[e,f]{Tom Bell}

\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
\affil[b]{Department of Environmental Sciences, University of Virginia}
\affil[c]{Department of Geography, University of California, Los Angeles}
\affil[d]{Marine Biological Association of the United Kingdom}
\affil[e]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
\affil[f]{Earth Research Institute, University of California, Santa Barbara}
\affil[*]{Corresponding author, reuman@ku.edu}

\begin{document}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=

#place R code here for loading in necessary variables from Results, Theory
resloc<-"../Results/Theory/"
N_A<-readRDS(paste0(resloc,"Params_A_N.rds"))
ns_A<-readRDS(paste0(resloc,"Params_A_ns.rds"))
p1_A<-readRDS(paste0(resloc,"Params_A_p1.rds"))
p2_A<-readRDS(paste0(resloc,"Params_A_p2.rds"))
b_A<-readRDS(paste0(resloc,"Params_A_b.rds"))

N_B<-readRDS(paste0(resloc,"Params_B_N.rds"))
ns_B<-readRDS(paste0(resloc,"Params_B_ns.rds"))
p1_B<-readRDS(paste0(resloc,"Params_B_p1.rds"))
p2_B<-readRDS(paste0(resloc,"Params_B_p2.rds"))
b_B<-readRDS(paste0(resloc,"Params_B_b.rds"))

N_C1<-readRDS(paste0(resloc,"Params_C1_N.rds"))
ns_C1<-readRDS(paste0(resloc,"Params_C1_ns.rds"))
p1_C1<-readRDS(paste0(resloc,"Params_C1_p1.rds"))
p2_C1<-readRDS(paste0(resloc,"Params_C1_p2.rds"))
b_C1<-readRDS(paste0(resloc,"Params_C1_b.rds"))

N_C2<-readRDS(paste0(resloc,"Params_C2_N.rds"))
ns_C2<-readRDS(paste0(resloc,"Params_C2_ns.rds"))
p1_C2<-readRDS(paste0(resloc,"Params_C2_p1.rds"))
p2_C2<-readRDS(paste0(resloc,"Params_C2_p2.rds"))
b_C2<-readRDS(paste0(resloc,"Params_C2_b.rds"))

resloc<-"../Results/Kelp_DataAfterSuperBasicCleaning/"
kelp_SB<-readRDS(paste0(resloc,"Kelp_Quarterly_CleanedSuperBasic.Rds"))
resloc<-"../Results/Kelp_DataAfterMoreInvolvedCleaning/"
kelp_MI<-readRDS(paste0(resloc,"Kelp_Quarterly_CleanedMoreInvolved.Rds"))

resloc<-"../Results/Kelp_SpectralDecompResults/"
CCA1l4<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag4.Rds"))
CCA2l4<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag4.Rds"))
SBl4<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag4.Rds"))
CCA1l8<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag8.Rds"))
CCA2l8<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag8.Rds"))
SBl8<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag8.Rds"))
CCA1l12<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag12.Rds"))
CCA2l12<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag12.Rds"))
SBl12<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag12.Rds"))
@

\maketitle

\tableofcontents
%\listoftables
\listoffigures

\section{Measuring synchrony} \label{Sect:measuring_sync}

If $w=(w_1,\ldots,w_N)$ is an $N$-dimensional stationary stochastic process, where the index $i=1,\ldots,N$ corresponds
to sampling locations, then we measure the synchrony of $w$ as 
\begin{linenomath}
\begin{equation}
\rho_{ww} = \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j},
\end{equation}
\end{linenomath}
where $S_{ww}$ is the spectral matrix of the process $w$. This is a function of frequency or timescale because
$S_{w_i w_j}$ is a function of frequency or timescale.
To see that $\rho_{ww}$ is a real-valued function, note that $\conj{S_{ww}}=S_{ww}^\tau$, where the superscript 
$\tau$ denotes matrix transpose [\cite{Brillinger2001}, theorem 2.5.1 on p. 24].
Therefore 
\begin{linenomath}
\begin{align}
\conj{\rho_{ww}} &= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \conj{S_{w_i w_j}} \\
&= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_j w_i} \\
&= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j} \\
&= \rho_{ww}.
\end{align} 
\end{linenomath}
The quantity $\rho_{ww}$ makes sense as a measure of synchrony in part because 
\begin{linenomath}
\begin{align}
\rho_{ww} &= \Rp(\rho_{ww}) \\
&= \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} \Rp(S_{w_i w_j}).
\end{align}
\end{linenomath}
The quantity $\Rp(S_{w_i w_j})$ is the cospectrum of the component processes $w_i$ and $w_j$, 
and an appropriate integral across frequencies of the cospectrum equals the covariance, 
$\cov(w_i,w_j)$ \citep{Brillinger2001}.
Thus an integral of $\rho_{ww}$ across frequencies equals 
\begin{linenomath}
\begin{equation}
\left( \frac{1}{N^2-N} \right) \sum_{i \neq j} \cov(w_i,w_j),
\end{equation}
\end{linenomath}
which is a standard quantity that can be used to measure non-frequency-specific synchrony.
Thus $\rho_{ww}$ is a frequency-specific generalization of a classic, non-frequency-specific 
measure of synchrony.

If $\epsilon^{(1)}=(\epsilon^{(1)}_1,\ldots,\epsilon^{(1)}_N)$ and 
$\epsilon^{(2)}=(\epsilon^{(2)}_1,\ldots,\epsilon^{(2)}_N)$
are two $N$-dimensional stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
(or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
\begin{linenomath}
\begin{equation}
\rho_{\epsilon^{(1)} \epsilon^{(2)}} = \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}.
\end{equation}
\end{linenomath}
This is a complex-valued function of frequency or timescale. Why is this interpretable as cross-variable 
spatial synchrony? A quantity representing spatial synchrony should compare values of the processes 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ in one location to values in another location, since spatial synchrony
is about relationships between distinct locations; hence our sum is over $i \neq j$. 
A quantity representing cross-variable synchrony should 
make comparisons between two variables, as opposed to comparisons between the same variable in different locations;
hence we use the cross spectrum between variable $1$ in location $i$ ($\epsilon^{(1)}_i$) and variable 
$2$ in location $j$ ($\epsilon^{(2)}_j$). 
Finally, a quantity representing cross-variable synchrony needs to
include the possibility of time delays in the relationship between the two variables. For instance, 
if $\epsilon^{(2)}_j$ is a time-lagged version of $\epsilon^{(1)}_i$, but the effects of 
$\epsilon^{(2)}_j$ on a biotic process $w_j$ are immediate, whereas the effects of $\epsilon^{(1)}_i$
on $w_i$ are delayed by the same lag, then the effects of $\epsilon^{(1)}_i$ on $w_i$ will be correlated
with the effects of $\epsilon^{(2)}_j$ on $w_j$, i.e., this combination of properties of $\epsilon^{(1)}$
and $\epsilon^{(2)}$ will produce synchrony, in the classic sense of correlation, between the biotic
processes $w_i$ and $w_j$.
Thus we use the cross spectrum, $S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}$, for which the phase angle at 
frequency $f$ measures the time delay between frequency-$f$ oscillations of $\epsilon^{(1)}_i$ and 
frequency-$f$ oscillations of $\epsilon^{(2)}_j$.

\section{Examples of cross synchrony}\label{sect:examples_cross_sync}

The following examples should help clarify the properties of our definition of cross synchrony.
For pedagogical clarity, we start with a white-noise example that does not illustrate the full value of the
idea of cross synchrony, but which is simple; we subsequently move on to an example which is more
complex but also more illustrative. These examples are also used in the theoretical case studies (main text).

For the simple example, to be used in theoretical case studies A and B (Methods), we 
assume $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are drawn, independently for each $t$, from a
$2N$-dimensional multivariate normal distribution
with mean $(0,...,0)$ and covariance matrix $\Sigma$, where the entries $\Sigma_{ij}$ of 
$\Sigma$ are as follows. For $i=j$, $\Sigma_{ij}=1$. For $i \neq j$ and $1 \leq i,j \leq N$, 
$\Sigma_{ij}$ equals a parameter $v_{11}$. For $i \neq j$ and $N+1 \leq i,j \leq 2N$, 
$\Sigma_{ij}$ equals a second parameter $v_{22}$. The remaining entries of $\Sigma$ equal a 
third parameter, $v_{12}$. The parameters $v_{11}$, $v_{22}$ and $v_{12}$ must be chosen to yield
a positive definite matrix $\Sigma$.

Now, having specified the noise process $(\epsilon^{(1)},\epsilon^{(2)})$, we derive the quantities 
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$.
For $a$ and $b$ equal to $1$ or $2$, the formula 
$S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = \sum_{u=-\infty}^\infty c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)
\exp(-2 \pi \iota f u)$, where $c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$,
follows from definitions of the spectrum and cross spectrum, given, for instance, in 2.5.2 on p. 23 of
\cite{Brillinger2001}. The quantity 
$c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$ is independent of 
$t$ for any stationary process $(\epsilon^{(1)},\epsilon^{(2)})$, and in particular for the temporally iid process that 
we assumed to operate here. Furthermore, because 
$\epsilon_i^{(a)}(t+u)$ and $\epsilon_j^{(b)}(t)$ were assumed independent for $u \neq 0$, we have 
\begin{linenomath}
\begin{equation}
S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(0) =
\begin{cases}
1 & \text{if}\ a=b\ \text{and}\ i=j \\
v_{11} & \text{if}\ a=b=1\ \text{and}\ i \neq j \\
v_{22} & \text{if}\ a=b=2\ \text{and}\ i \neq j \\
v_{12} & \text{if}\ a \neq b.
\end{cases}
\end{equation}
\end{linenomath}
Thus the spectra and cross spectra in this example are constant. It
is then easy to see that $\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$, based on the definitions of these 
quantities given previously. For theoretical case studies A and B in the main text, we used
$N=\Sexpr{N_A}$, $v_{11}=v_{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.

These results make intuitive sense, for a white-noise process, for the following reasons.
First, $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$ and $\rho_{\epsilon^{(2)} \epsilon^{(2)}}$,
which are automatically real-valued (SI section \ref{Sect:measuring_sync}), should be flat for a white noise
process, i.e., independent of timescale, because spectral quantities are flat 
for time series with no temporal dependencies.  For the same reason, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
should be flat. It makes sense
that cross synchrony in this example is real-valued because a complex-valued 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ would imply lagged or phase-delayed relationships between 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$, not possible with a white-noise process.

For a more illustrative example, used in case study C (Methods), the noise process 
$(\epsilon^{(1)},\epsilon^{(2)})$ is generated via a three-step process. 
First, assume $(\tilde{\zeta}^{(1)}(t),\tilde{\zeta}^{(2)}(t))$ 
are drawn, independently for each $t$, from a $2N$-dimensional multivariate normal distribution with mean $(0,\ldots,0)$
and covariance matrix $\Psi$, where the entries $\Psi_{ij}$ of $\Psi$ are as follows. 
For $i=j$, $\Psi_{ij}=1$. For $i \neq j$ and $1 \leq i,j \leq N$, 
$\Psi_{ij}$ equals a parameter $w_{11}$. For $i \neq j$ and $N+1 \leq i,j \leq 2N$, 
$\Psi_{ij}$ equals a second parameter $w_{22}$. The remaining entries of $\Psi$ equal a 
third parameter, $w_{12}$. The parameters $w_{11}$, $w_{22}$ and $w_{12}$ must again be chosen to yield
a positive definite matrix. Second, let $\zeta^{(1)}(t)=\tilde{\zeta}^{(1)}(t-l_1)$ and let
$\zeta^{(2)}(t)=\tilde{\zeta}^{(2)}(t-l_2)$ for fixed integers $l_1$ and $l_2$. Finally,
let $\epsilon_i^{(a)}(t)=c_1 \epsilon_i^{(a)}(t-1)+c_2 \epsilon_i^{(a)}(t-2)+\zeta_i^{(a)}(t)$
for $i=1,\ldots,N$ and $a=1,2$. The second step above ensures that $\epsilon^{(1)}$ and $\epsilon^{(2)}$
are related to each other in a time-delayed manner. The third step ensures that 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ can oscillate with a dominant periodic component, depending 
on the values of $c_1$ and $c_2$ used: the autoregressive order-two process used
can produce dynamics with a peaked spectrum, as we will see below (see also theorem \ref{thm:useful_thm} below).

To compute $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ for this example, we first compute 
$S_{\epsilon^{(1)}\epsilon^{(1)}}$, $S_{\epsilon^{(2)}\epsilon^{(2)}}$ and
$S_{\epsilon^{(1)}\epsilon^{(2)}}$ using theorem \ref{thm:useful_thm} below.  
Letting $\epsilon=(\epsilon^{(1)},\epsilon^{(2)})$ and $\zeta(t)=(\zeta^{(1)},\zeta^{(2)})$, we 
have $\epsilon_i(t)=c_1 \epsilon_i(t-1)+c_2 \epsilon_i(t-2) + \zeta_i(t)$ for $i=1,\ldots,2N$. We then
apply theorem \ref{thm:useful_thm} to get that $S_{\epsilon \epsilon}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta \zeta}$.
Considering upper-left $N \times N$ sub-matrices gives 
$S_{\epsilon^{(1)} \epsilon^{(1)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(1)} \zeta^{(1)}}$.
Considering lower-right $N \times N$ sub-matrices gives 
$S_{\epsilon^{(2)} \epsilon^{(2)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(2)} \zeta^{(2)}}$.
Considering upper-right $N \times N$ sub-matrices gives 
$S_{\epsilon^{(1)} \epsilon^{(2)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(1)} \zeta^{(2)}}$.
Because the $\zeta^{(1)}(t)$ are iid for distinct times $t$, we can again use the definitions of the spectrum and the
cross spectrum and reasoning similar to that used above for the previous examples to show that $S_{\zeta^{(1)} \zeta^{(1)}}$
equals the upper-left $N \times N$ sub-matrix of $\Psi$, which is a matrix with $1$ in every diagonal entry and
$w_{11}$ in all off-diagonal entries. Likewise, $S_{\zeta^{(2)} \zeta^{(2)}}$ equals the lower-right 
$N \times N$ sub-matrix of $\Psi$, which has $1$ in all diagonal entries and $w_{22}$ in all off-diagonal entries.
To get $S_{\zeta^{(1)} \zeta^{(2)}}$, we write 
$S_{\zeta_i^{(1)}\zeta_j^{(2)}}(f)=\sum_{u=-\infty}^{\infty} c_{\zeta_i^{(1)}\zeta_j^{(2)}}(u) \exp(-2 \pi \iota f u)$.
Here, 
\begin{linenomath}
\begin{align}
c_{\zeta_i^{(1)}\zeta_j^{(2)}}(u) &= \cov(\zeta_i^{(1)}(t+u),\zeta_j^{(2)}(t)) \\
&= \cov(\tilde{\zeta}_i^{(1)}(t+u-l_1),\tilde{\zeta}_j^{(2)}(t-l_2)) \\
&= 
\begin{cases}
w_{12} & \text{if}\ u-l_1=-l_2 \\
0 & \text{otherwise}.
\end{cases}
\end{align}
\end{linenomath}
So $S_{\zeta_i^{(1)}\zeta_j^{(2)}}=w_{12} \exp(-2 \pi \iota f (l_1 - l_2))$. It then follows 
straightforwardly that $\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
Fig. \ref{fig:rhosForC} displays these quantities for the two parameterizations
of $(\epsilon^{(1)},\epsilon^{(2)})$ that we use for theoretical case study C.

We now explore why the results above for $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ make sense for the noise processes we described, particularly for the sets of parameters 
used for theoretical case study C in the main text (see Fig. \ref{fig:rhosForC}). 
The pre-factor $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ in all these expressions reflects the autoregressive 
manner in which the noises $\epsilon$ were defined. Autoregressive models of order $2$ are well known to be capable of
producing oscillations for certain values of the parameters. For instance, for case study C, we 
use $c_1=\Sexpr{ns_C1[1]}$ and $c_2=\Sexpr{round(ns_C1[2],4)}$, values
which cause the expression $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ to have a single peak at
frequency $0.25$ cycles per sampling interval. It makes sense that the synchrony quantities
$\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ for periodically oscillating quantities should also have a peak at the 
timescale of oscillation. 
The term $\exp(-2 \pi \iota f (l_1-l_2))$ in the expression for 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ reflects that fact that cross variable synchrony for this example is lagged
by an amount $l_1-l_2$. The lack of a similar term in the expressions for $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$
or $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ reflects that fact that there are no lags built into within-variable
synchrony in this example.

\begin{figure}
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01_rhos.pdf}\\
\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02_rhos.pdf}
\caption[Spectral quantities for noise used in case study C]{The quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the noise
described in SI section \ref{sect:examples_cross_sync} and used for theoretical case study C. 
These are complex-valued functions of timescale; their magnitude is displayed on vertical axes and
their phase is displayed in color. Two alternative sets of parameters
were used for panels a-c and d-f, respectively, and are also used in the two scenarios
considered for case study C. Parameters were:
$N=\Sexpr{N_C1}$, $c_1=\Sexpr{ns_C1[1]}$, $c_2=\Sexpr{round(ns_C1[2],4)}$, 
$w_{11}=w{22}=\Sexpr{ns_C1[5]}$ and $w_{12}=\Sexpr{ns_C1[7]}$ for both sets of panels;
$l_1=\Sexpr{ns_C1[3]}$ and $l_2=\Sexpr{ns_C1[4]}$ for panels a-c; and 
$l_1=\Sexpr{ns_C2[3]}$ and $l_2=\Sexpr{ns_C2[4]}$ for panels d-f.
}\label{fig:rhosForC}
\end{figure}

\begin{thm}\label{thm:useful_thm}
If $\zeta(t)$ is an $N$-dimensional second-order stationary ergodic stochastic process with 
component expected values $0$, if
$\epsilon_i(t)=c_1 \epsilon_i(t-1)+\cdots+c_n \epsilon_i(t-n)+q_0 \zeta_i(t)+\cdots+q_m \zeta_i(t-m)$ for $i=1,\ldots,N$, 
with $c_n \neq 0$ and $q_m \neq 0$, and if the complex roots of the polynomial
$1-c_1 z - c_2 z^2 - \cdots c_n z^n = 0$ have modulus greater than $1$, then 
$S_{\epsilon \epsilon}(f) = \left| \frac{\gamma}{1-\lambda} \right|^2 S_{\zeta \zeta}(f)$ where $\gamma=q_0+q_1 \mu+\cdots+q_m \mu^m$,
$\lambda=c_1 \mu+c_2 \mu^2+\cdots+c_n \mu^n$, and $\mu=\exp(-2\pi \iota f)$, $\iota$ is the imaginary unit, and 
$f$ is frequency in cycles per time step.
\end{thm}
\begin{proof}
This theorem is identical to theorem 2 in SI section S1.5 of \cite{Anderson2021}, where a proof is also provided. 
\end{proof}

\section{Model specification, full details}\label{sect:model_spec}

The population model is specified starting with the equation
\begin{linenomath}
\begin{align}
w_i(t) &= b_1 w_i(t-1)+ \cdots + b_n w_i(t-n) \label{main_model_1} \\ 
&+ p_0^{(1)} \epsilon_i^{(1)}(t)+ \cdots + p_{m_1}^{(1)} \epsilon_i^{(1)}(t-m_1) \label{main_model_2} \\
&+ p_0^{(2)} \epsilon_i^{(2)}(t)+ \cdots + p_{m_2}^{(2)} \epsilon_i^{(2)}(t-m_2) \label{main_model_3} \\
&+ \delta_i(t), \label{main_model_4}
\end{align}
\end{linenomath}
where $i=1,\ldots,N$ indexes habitat patches or sampling locations, $w_i(t)$ is a measure of the 
population in location $i$ at time $t$, and $\epsilon^{(1)}=(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)})$,
$\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$ are
environmental processes influencing the populations. Note that this model assumes from the outset 
a homogeneous influence of the environmental variables on the populations in the different
sampling locations, i.e., the parameters $p$ are spatially homogeneous. 
%We also assume spatial homogeneity of the AR coefficients, need to add a mention of that
We also assume that the 
combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process \citep{Brillinger2001}
with the expected values of its component processes equal to $0$. Finally, we assume the complex roots of the
polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, an assumption needed for the 
existence of a stationary stochastic process solution for the model and for invertability of
the filter $\mcB$ which we will define below. See \cite{Reinsel1997,Shumway2000,Brillinger2001}
for background on the statistical theory of stationary stochastic processes.

Letting $I_N$ denote the $N \times N$ identity matrix and letting $B$ denote the backshift operator, 
we define the linear filters 
\begin{linenomath}
\begin{align}
\mcP^{(1)} &= p_0^{(1)} I_N + p_1^{(1)} I_N B + \cdots + p_{m_1}^{(1)} I_N B^{m_1}, \\
\mcP^{(2)} &= p_0^{(2)} I_N + p_1^{(2)} I_N B + \cdots + p_{m_2}^{(2)} I_N B^{m_2}, \\
\mcB &= I_N - b_1 I_N B - b_2 I_N B^2 - \cdots b_n I_N B^n.
\end{align}
\end{linenomath}
Using the language of filters and stochastic processes, the original model 
(\ref{main_model_1}-\ref{main_model_4}) can then be written as
\begin{linenomath}
\begin{equation}
\mcB w = \mcP^{(1)} \epsilon^{(1)} +\mcP^{(2)} \epsilon^{(2)} +\delta.\label{main_model_stochproc}
\end{equation}
\end{linenomath}
Writing the model in the language of stochastic processes and linear filters makes it straightforward
to calculate the spectral matrix, $S_{ww}$, of the stochastic process, $w$, which we do in section \ref{sect:derivation}.

\section{Derivation of the main equation}\label{sect:derivation}

Computing the spectral matrix of both sides of the model (\ref{main_model_stochproc}) gives
\begin{linenomath}
\begin{align}
T(\mcB)S_{ww} T(\mcB)^* &= T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(1)}} T(\mcP^{(1)})^* \label{spectmat_eq1_t1}\\
  &+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
  &+ S_{\delta \delta} \\
  &+ T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(2)}} T(\mcP^{(2)})^*+T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(1)}} T(\mcP^{(1)})^* \\
  &+ T(\mcP^{(1)}) S_{\epsilon^{(1)}\delta} + S_{\delta \epsilon^{(1)}} T(\mcP^{(1)})^* \\
  &+ T(\mcP^{(2)}) S_{\epsilon^{(2)}\delta} + S_{\delta \epsilon^{(2)}} T(\mcP^{(2)})^*, \label{spectmat_eq1_t6}
\end{align}
\end{linenomath}
where $T(\mcB)$, $T(\mcP^{(1)})$ and $T(\mcP^{(2)})$ are the tranfer functions of the filters
$\mcB$, $\mcP^{(1)}$ and $\mcP^{(2)}$, respectively, $S_{ww}$ is the spectral matrix of the process $w$,
$S_{\epsilon^{(i)} \epsilon^{(i)}}$ is the spectral matrix of the process $\epsilon^{(i)}$, $S_{\delta \delta}$
is the spectral matrix of the process $\delta$,  $S_{\epsilon^{(i)} \epsilon^{(j)}}$ is the cross spectral
matrix of the processes $\epsilon^{(i)}$ and $\epsilon^{(j)}$, $S_{\epsilon^{(i)} \delta}$ is the cross 
spectral matrix of the processes $\epsilon^{(i)}$ and $\delta$, $S_{\delta \epsilon^{(i)}}$ is the cross
spectral matrix of the processes $\delta$ and $\epsilon^{(i)}$, and the superscript $*$ denotes
the conjugate transpose of a matrix.

Letting $\mu=\exp(-2\pi \iota f)$ where $\iota$ is the imaginary unit and 
$f$ is frequency in units of cycles per time step, we have
\begin{linenomath}
\begin{align}
T(\mcB) &= (1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n) I_N \\
T(\mcP^{(1)}) &= (p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1}) I_N \\
T(\mcP^{(2)}) &= (p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}) I_N.
\end{align}
\end{linenomath}
\noindent These are all scalar matrices, so (\ref{spectmat_eq1_t1}-\ref{spectmat_eq1_t6}) simplifies to
\begin{linenomath}
\begin{align}
S_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} S_{\epsilon^{(1)} \epsilon^{(1)}} \label{Sww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} S_{\epsilon^{(2)} \epsilon^{(2)}} \\
&+ \frac{1}{|f_{\mcB}|^2} S_{\delta \delta}\\
&+ \frac{1}{|f_{\mcB}|^2} \left[  f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}  S_{\epsilon^{(1)} \epsilon^{(2)}} + 
f_{\mcP^{(2)}} \conj{f_{\mcP^{(1)}}} S_{\epsilon^{(2)} \epsilon^{(1)}}\right] \label{Sww_T4}\\
&+ \frac{1}{|f_{\mcB}|^2} \left[f_{\mcP^{(1)}} S_{\epsilon^{(1)} \delta} + \conj{f_{\mcP^{(1)}}}  S_{\delta \epsilon^{(1)}}\right]\\
&+ \frac{1}{|f_{\mcB}|^2} \left[ f_{\mcP^{(2)}} S_{\epsilon^{(2)} \delta} + \conj{f_{\mcP^{(2)}}}  S_{\delta \epsilon^{(2)}} \right], \label{Sww_T6}
\end{align}
\end{linenomath}
where 
\begin{linenomath}
\begin{align}
f_{\mcB} &= 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n \\
f_{\mcP^{(1)}} &= p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1} \\
f_{\mcP^{(2)}} &= p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}.
\end{align}
\end{linenomath}
We here made use of the assumption that the complex roots of the
polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, so that $f_{\mcB}$ is non-zero.

Now, averaging the off-diagonal entries of both sides of (\ref{Sww_T1}-\ref{Sww_T6}) gives
\begin{linenomath}
\begin{align}
\left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{w_i w_j} &= 
\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
 &+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}} \\
 &+ \frac{1}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} 
  \Rp \left( f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}} \right) \label{SumSww_T4}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right)  \sum_{i \neq j} \Rp \left( f_{\mcP^{(1)}}  
 S_{\epsilon_i^{(1)} \delta_j}\right)  \label{SumSww_T5}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} \Rp \left( f_{\mcP^{(2)}}    
 S_{\epsilon_i^{(2)} \delta_j}\right), \label{SumSww_T6}
\end{align}
\end{linenomath}
\noindent where $\Rp(\cdot)$ denotes the real part, and where (\ref{SumSww_T4}-\ref{SumSww_T6})
follows from two observations: that the two summands in the brackets of each of the terms 
(\ref{Sww_T4}-\ref{Sww_T6}) are conjugate-transpose pairs; and that, for any complex matrix, $A$,
$\sum_{i \neq j} (A_{ij} +\conj{A_{ji}}) = \sum_{i \neq j} (A_{ij} +\conj{A_{ij}}) = 2 \sum_{i \neq j} \Rp (A_{ij})$.
Now, making use of the definitions of section \ref{Sect:measuring_sync}, this becomes
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}
+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}
+ \frac{1}{|f_{\mcB}|^2} \rho_{\delta \delta} \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(1)}} \rho_{\epsilon^{(1)} \delta} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right),
\end{align}
\end{linenomath}
which expresses population synchrony in terms of noise synchrony and cross synchrony.
If the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ were measured and the process $\delta$ represents
the aggregate effects of unknown processes, then we can write
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}} \label{rhoww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}} \label{rhoww_T2}\\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \label{rhoww_Tint} \\
&+ \text{other contributions}.
\end{align}
\end{linenomath}
The term (\ref{rhoww_T1}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(1)}$.
The term (\ref{rhoww_T2}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(2)}$.
The term (\ref{rhoww_Tint}) is contributions to the synchrony of $w$ due to interactions between the Moran effects
of $\epsilon^{(1)}$ and $\epsilon^{(2)}$. The ``other contributions'' may include direct Moran effects of other,
unmeasured environmental factors, as well as interactions between these effects and Moran effects of the 
measured factors $\epsilon^{(1)}$ and $\epsilon^{(2)}$.

\section{Details of data and data cleaning}\label{sect:data}

Floating surface canopy of kelp was measured by remote sensing, with details and
validation reported by \cite{Cavanaugh2011} and \cite{Bell2020, Bell2021}. Multispectral Landsat
imagery of $30$m resolution was used. Quarterly mean biomass values and spatial extent of canopy  
within $500$m coastline segments (henceforth called ``locations'')
were computed by averaging. Quarters were: Q1, January-March; Q2, April-June; 
Q3, July-September; Q4, October-December. 
Data were thusly available from $\Sexpr{dim(kelp_SB)[1]}$ locations spanning 
approximately from Monterey Bay to San Diego. 
We then chose locations for analysis for which kelp was ``persistent'', in this 
context meaning absent or missing data for no more than three years; this was to focus on 
population dynamics of kelp rather than extinction-recolonization dynamics, which occur in some 
locations and which can relate to sediment dynamics on the ocean floor and other factors
not currently of primary interest. Of the original $\Sexpr{dim(kelp_SB)[1]}$ locations,
$\Sexpr{dim(kelp_MI)[1]}$ ($\Sexpr{100*round(dim(kelp_MI)[1]/dim(kelp_SB)[1],2)}$\%) 
were deemed persistent according to this criterion. Missing values in the remaining 
time series were then filled with seasonal medians. The biomass and spatial extent time series
for each location were then used together to produce one time series estimating biomass
density per unit useable habitat at the site, as follows: every value of 
the biomass time series was divided by the $90^{\text{th}}$ quantile of the distribution
of the spatial extent time series at the site. The $90^{\text{th}}$ quantile was used instead
of the maximum value of the spatial extent time series because maximum statistics have high variance. 
This procedure served to normalize total biomass by an estimate
of total useable habitat at a site. Kelp establish on rocky substrate, and areas with sandy or
sedimented substrate are not useable kelp habitat. All time series were then linearly detrended,
and only those time series were utilized which were located within the three regions pictured
in Fig. \ref{MT-fig:map} in the main text. Regions were selected to represent distinct dynamical
patterns known to occur in central and southern CA, and also to have another the same number
of time series per selected region.

Wave nutrient data were also available for each quarter at each location. Wave data were derived 
by applying a swell propagation model with bouy-based wave measurements and a validation 
techniques \citep{Wingeart2001,Hansen2009,Bell2015,OReilly2016}. Data on nitrates 
were produced \citep{Snyder2020} by exploiting an established empirical 
relationship between surface nitrate concentration
and sea surface temperature (SST) in coastal CA \citep{Zimmerman1984,Palacios2013,Snyder2020}
and SST data derived from Advanced Very High Resolution Radiometer satellite imagery. 
Wave and nutrient time series in each location were again linearly detrended before use.

\section{Details of the theoretical case studies}\label{sect:details_case_studies}

The noise processes $(\epsilon^{(1)},\epsilon^{(2)})$ used 
in case studies A, B and C, which were only partly specified in the main text, are fully specified in 
SI section \ref{sect:examples_cross_sync}. 
We next derive, for our case studies, the values of the terms that appear in (\ref{MT-rhoww_T1to3}) of the main text.
For case study A, parameter specifications presented in Methods of the main text straightforwardly imply that
$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_1^{(1)} \mu$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. In 
SI section \ref{sect:examples_cross_sync}, we computed that 
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.

For case study B, details presented in Methods of the main text imply that 
$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
Because the noise process for B is taken to be the same as that for A, we again have 
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$ and 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.

For case study C, again using details from Methods, $f_{\mcB}=1-b_1 \mu$, 
$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$. 
In SI section \ref{sect:examples_cross_sync}, we computed that
$\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.

\section{Supporting results for kelp}

\begin{figure}[!h]
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_MainPlot_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_MainPlot_ForMS.pdf}
\caption[Extension of main text Fig. \ref{MT-fig:SummaryKelpFig}]{Same format as Fig. \ref{MT-fig:SummaryKelpFig} of the main text, but including not only results
for the CCA1 region (a-c) and the SB region (d-f), but also results from the 
CCA2 region (g-i).}\label{fig:SummaryKelpFig_SM_lag4}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_MainPlot_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_MainPlot_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_MainPlot_ForMS.pdf}
\caption[Like Fig. \ref{fig:SummaryKelpFig_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:SummaryKelpFig_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:SummaryKelpFig_SM_lag4}.}\label{fig:SummaryKelpFig_SM_lag8}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_MainPlot_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_MainPlot_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_MainPlot_ForMS.pdf}
\caption[Like Fig. \ref{fig:SummaryKelpFig_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:SummaryKelpFig_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:SummaryKelpFig_SM_lag4}.}\label{fig:SummaryKelpFig_SM_lag12}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\caption[Extension of main text Fig. \ref{MT-fig:directMoran}]{Same format as Fig. \ref{MT-fig:directMoran} of the main text, but including not only results
for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
CCA2 region (ee-ss).}\label{fig:directMoran_SM_lag4}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_ExplainDirectMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_ExplainDirectMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_ExplainDirectMoran_ForMS.pdf}
\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:directMoran_SM_lag4}.}\label{fig:directMoran_SM_lag8}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_ExplainDirectMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_ExplainDirectMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_ExplainDirectMoran_ForMS.pdf}
\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:directMoran_SM_lag4}.}\label{fig:directMoran_SM_lag12}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
\caption[Extension of main text Fig. \ref{MT-fig:indirectMoran}]{Same format as Fig. \ref{MT-fig:indirectMoran} of the main text, but including not only results
for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
CCA2 region (ee-ss).}\label{fig:indirectMoran_SM_lag4}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_ExplainInteractingMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_ExplainInteractingMoran_ForMS.pdf} 
\caption[Like Fig. \ref{fig:indirectMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:indirectMoran_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:indirectMoran_SM_lag4}.}\label{fig:indirectMoran_SM_lag8}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_ExplainInteractingMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_ExplainInteractingMoran_ForMS.pdf} 
\caption[Like Fig. \ref{fig:indirectMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:indirectMoran_SM_lag4}, but based on a model that considered 
potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
the same as for Fig. \ref{fig:indirectMoran_SM_lag4}.}\label{fig:indirectMoran_SM_lag12}
\end{figure}

\clearpage

\begin{table}[!h]
\caption{Regression coefficients used to parameters (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each 
of our three regions. All values rounded to two decimal places.}\label{tab:regcoefs}
\begin{center}
\begin{tabular}{r|l|l|l|l|l}
 & $p_0^{(1)}$ & $p_1^{(1)}$ & $p_0^{(2)}$ & $b_1,b_2,\ldots,b_n$ & $R^2$\\
 \hline
%n=4
CCA1, $n=4$ & \Sexpr{round(unname(CCA1l4[6]),2)} & \Sexpr{round(unname(CCA1l4[7]),2)} & \Sexpr{round(unname(CCA1l4[8]),2)} & \Sexpr{paste0(round(unname(CCA1l4[2:5]),2),collapse=",")} & \Sexpr{round(unname(CCA1l4[1]),2)} \\
CCA2, $n=4$ & \Sexpr{round(unname(CCA2l4[6]),2)} & \Sexpr{round(unname(CCA2l4[7]),2)} & \Sexpr{round(unname(CCA2l4[8]),2)} & \Sexpr{paste0(round(unname(CCA2l4[2:5]),2),collapse=",")} & \Sexpr{round(unname(CCA2l4[1]),2)} \\
SB, $n=4$ & \Sexpr{round(unname(SBl4[6]),2)} & \Sexpr{round(unname(SBl4[7]),2)} & \Sexpr{round(unname(SBl4[8]),2)} & \Sexpr{paste0(round(unname(SBl4[2:5]),2),collapse=",")} & \Sexpr{round(unname(SBl4[1]),2)} \\
%n=8
CCA1, $n=8$ & \Sexpr{round(unname(CCA1l8[10]),2)} & \Sexpr{round(unname(CCA1l8[11]),2)} & \Sexpr{round(unname(CCA1l8[12]),2)} & \Sexpr{paste0(round(unname(CCA1l8[2:9]),2),collapse=",")} & \Sexpr{round(unname(CCA1l8[1]),2)} \\
CCA2, $n=8$ & \Sexpr{round(unname(CCA2l8[10]),2)} & \Sexpr{round(unname(CCA2l8[11]),2)} & \Sexpr{round(unname(CCA2l8[12]),2)} & \Sexpr{paste0(round(unname(CCA2l8[2:9]),2),collapse=",")} & \Sexpr{round(unname(CCA2l8[1]),2)} \\
SB, $n=8$ & \Sexpr{round(unname(SBl8[10]),2)} & \Sexpr{round(unname(SBl8[11]),2)} & \Sexpr{round(unname(SBl8[12]),2)} & \Sexpr{paste0(round(unname(SBl8[2:9]),2),collapse=",")} & \Sexpr{round(unname(SBl8[1]),2)} \\
%n=12
CCA1, $n=12$ & \Sexpr{round(unname(CCA1l12[14]),2)} & \Sexpr{round(unname(CCA1l12[15]),2)} & \Sexpr{round(unname(CCA1l12[16]),2)} & \Sexpr{paste0(round(unname(CCA1l12[2:13]),2),collapse=",")} & \Sexpr{round(unname(CCA1l12[1]),2)} \\
CCA2, $n=12$ & \Sexpr{round(unname(CCA2l12[14]),2)} & \Sexpr{round(unname(CCA2l12[15]),2)} & \Sexpr{round(unname(CCA2l12[16]),2)} & \Sexpr{paste0(round(unname(CCA2l12[2:13]),2),collapse=",")} & \Sexpr{round(unname(CCA2l12[1]),2)} \\
SB, $n=12$ & \Sexpr{round(unname(SBl12[14]),2)} & \Sexpr{round(unname(SBl12[15]),2)} & \Sexpr{round(unname(SBl12[16]),2)} & \Sexpr{paste0(round(unname(SBl12[2:13]),2),collapse=",")} & \Sexpr{round(unname(SBl12[1]),2)} \\
\end{tabular}
\end{center}
\end{table}

\bibliographystyle{ecology_letters2}
\bibliography{refs}

\end{document}